---
title: "ENVS 193DS Homework 3"
date: "5/23/26" 
author: "Steven Le"
format: pdf 
execute:
  warning: false
  message: false
editor: source
include-in-header:
      text: |
        \usepackage{fvextra}
        \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

# Homework set up

```{r}

#read in packages
library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(ggplot2)
library(ggeffects)
library(flextable)
#stored kelp data as an object called kelp
kelp <- read_csv("temp-kelp.csv")
#stored personal data as an object called personal
personal <- read_csv("personal_data.csv")

```

# Problem 1. Giant kelp fronds 

## a. 

To determine the strength of the relationship between temperature and giant kelp frond elongation, you need to conduct a Pearson's correlation test and Spearman rank correlation test.  A Pearson's correlation test is a parametric test that analyzes a linear relationship between continuous variables that are normally distributed (debatable).  On the other hand, a Spearman rank correlation test is the non-parametric version that looks as a monotonic relationship between variables that have independently observed samples.  

## b.

```{r}

#base layer: ggplot
ggplot(data = kelp,
#x-axis: temperature, y-axis: kelp elongation rate       
       mapping = aes(x = temp_c,
                     y = kelp_elong)) + 
#first layer: geom_point
  geom_point(color = "darkgreen") + #changed color of points to dark green
#renamed x- and y-axes and added units  
  labs(x = "Temperature (\u00B0C)",
       y = "Kelp Elongation Rate (cm/day)") +
#changed theme from default  
  theme_minimal()

```

## c.

## QQ plot figure for temperature:

```{r}

#base layer: ggplot with starting data frame
ggplot(data = kelp,
#y-axis for QQ plot (no x-axis)
aes(sample = temp_c)) +
#first layer: QQ reference line colored red
geom_qq_line(color = "red") +
#second layer: QQ plot
geom_qq() 

```

## QQ plot figure for giant kelp frond elongation rate:

```{r}

#base layer: ggplot with starting data frame
ggplot(data = kelp,
#y-axis for QQ plot (no x-axis)
aes(sample = kelp_elong)) +
#first layer: QQ reference line colored red
geom_qq_line(color = "red") +
#second layer: QQ plot
geom_qq() 

```

## Evaluating Pearson's correlation test assumptions:

By making QQ plots for each of the two variables, we were able to interpret if the variables, respectively, are normally distributed.  Looking at both QQ plots, temperature (Celsius) and giant kelp frond elongation rate (cm/day) look normally distributed as most of the points roughly follow the reference line in each plot.  For the other three assumptions, there does seem to be a linear (negative) relationship between the variables based on the scatterplot, the variables are definitely continuous, and we can give the benefit of the doubt that the data were collected through independent observations.    

## Pearson's correlation test code chunk:

```{r}

#ran a correlation test between temperature and kelp elongation rate
cor.test(kelp$kelp_elong, kelp$temp_c,
#used Pearson's r correlation test as our method         
         method = "pearson")

```

## d.

To evaluate the strength of the relationship between temperature and giant kelp frond elongation rate, I used a Pearson's correlation test since each of the variables has a roughly normal distribution, the variables are continuous, and the sample size is large (n > 30).  We found a moderate negative relationship between temperature (Celsius) and giant kelp frond elongation rate (cm/day) (Pearson's r = -0.69, t(30) = -5.19, p < 0.0001, $\alpha$ = 0.05).

## e.



## f.



# Problem 2. Personal data

## a.



## b.



# Problem 3. Affective visualization

## a.



## b.



## c.



## d.



## e.



# Problem 4. Statistical critique

## a.



## b.



## c.



## d.


