---
title: "ENVS 193DS Homework 3"
date: "5/23/26" 
author: "Steven Le"
format: pdf 
execute:
  warning: false
  message: false
editor: source
include-in-header:
      text: |
        \usepackage{fvextra}
        \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

# Homework set up

```{r}

#read in packages
library(tidyverse)
library(here)
library(janitor)
library(readxl)
library(ggplot2)
library(ggeffects)
library(flextable)
#stored kelp data as an object called kelp
kelp <- read_csv("data/temp-kelp.csv")
#stored personal data as an object called personal
personal <- read_csv("data/personal_data.csv")

```

# Problem 1. Giant kelp fronds 

## a. 

To determine the strength of the relationship between temperature and giant kelp frond elongation, you need to conduct a Pearson's correlation test and Spearman rank correlation test.  A Pearson's correlation test is a parametric test that analyzes a linear relationship between continuous variables that are normally distributed (debatable).  On the other hand, a Spearman rank correlation test is the non-parametric version that looks as a monotonic relationship between variables that have independently observed samples.  

## b.

```{r}

#base layer: ggplot
ggplot(data = kelp,
#x-axis: temperature, y-axis: kelp elongation rate       
       mapping = aes(x = temp_c,
                     y = kelp_elong)) + 
#first layer: geom_point
  geom_point(color = "darkgreen") + #changed color of points to dark green
#renamed x- and y-axes and added units  
  labs(x = "Temperature (\u00B0C)",
       y = "Kelp Elongation Rate (cm/day)") +
#changed theme from default  
  theme_minimal()

```

## c.

## QQ plot figure for temperature:

```{r}

#base layer: ggplot with starting data frame
ggplot(data = kelp,
#y-axis for QQ plot (no x-axis)
aes(sample = temp_c)) +
#first layer: QQ reference line colored red
geom_qq_line(color = "red") +
#second layer: QQ plot
geom_qq() 

```

## QQ plot figure for giant kelp frond elongation rate:

```{r}

#base layer: ggplot with starting data frame
ggplot(data = kelp,
#y-axis for QQ plot (no x-axis)
aes(sample = kelp_elong)) +
#first layer: QQ reference line colored red
geom_qq_line(color = "red") +
#second layer: QQ plot
geom_qq() 

```

## Evaluating Pearson's correlation test assumptions:

By making QQ plots for each of the two variables, we were able to interpret if the variables, respectively, are normally distributed.  Looking at both QQ plots, temperature (Celsius) and giant kelp frond elongation rate (cm/day) look normally distributed as most of the points roughly follow the reference line in each plot.  For the other three assumptions, there does seem to be a linear (negative) relationship between the variables based on the scatterplot, the variables are definitely continuous, and we can give the benefit of the doubt that the data were collected through independent observations.    

## Pearson's correlation test code chunk:

```{r}

#ran a correlation test between temperature and kelp elongation rate
cor.test(kelp$kelp_elong, kelp$temp_c,
#used Pearson's r correlation test as our method         
         method = "pearson")

```

## d.

To evaluate the strength of the relationship between temperature and giant kelp frond elongation rate, I used a Pearson's correlation test since each of the variables has a roughly normal distribution, the variables are continuous, and the sample size is large (n > 30).  We found a moderate negative relationship between temperature (Celsius) and giant kelp frond elongation rate (cm/day) (Pearson's r = -0.69, t(30) = -5.19, p < 0.0001, $\alpha$ = 0.05).

## e.

Giant kelp seems to grow faster in cooler temperatures as our data seems to suggest that as the temperature (Celsius) increases, the elongation rate (cm/day) of giant kelp frond elongation rate slows down.  In other words, the highest elongation rates were found in the coolest temperatures recorded whereas the slowest elongatin rates were found in the highest temperatures recorded.  Therefore, giant kelp health is most likely better maintained in temperatures that are not as warm since it would help with their frond elongation to continue growing at a higher rate.   

## f.

```{r}

#ran a correlation test between temperature and kelp elongation rate
cor.test(kelp$kelp_elong, kelp$temp_c,
#used Spearman rank correlation test as our method         
         method = "spearman", 
#set exact equal to FALSE to allow approximation
exact = FALSE)

```

The Pearson's correlation and Spearman rank correlation tests would have led us to the same decision about the null hypothesis as the p-value is still less than our significance level of 0.05, which means we can reject the null hypothesis.  Therefore, there does seem to be a relationship between between temperature (Celsius) and giant kelp frond elongation rate (cm/day).  In addition, the Spearman rank correlation value was roughly -0.69, which was the same as our Pearson's correlation value, so there seems to be a a moderate negative relationship between temperature (Celsius) and giant kelp frond elongation rate (cm/day) (Spearman $\rho$ = -0.69, S = 9216, p < 0.0001, $\alpha$ = 0.05).   

# Problem 2. Personal data

## a.

```{r}

#created new object from personal
personal_clean <- personal |>
#cleaned column names    
  clean_names() |> 
#mutated the existing weather column  
  mutate(weather = case_when(
#filled in Sunny to replace sunny    
    weather == "sunny"  ~ "Sunny",
#filled in Cloudy to replace cloudy    
    weather == "cloudy" ~ "Cloudy",
#filled in Windy to replace windy    
    weather == "windy"  ~ "Windy",
#filled in Night to replace night    
    weather == "night"  ~ "Night"))

```

```{r}

#base layer: ggplot with x-axis
ggplot(data = personal_clean, 
#x-axis is weather and y-axis is frequency of refilling water bottle       
       mapping = aes(x = weather)) +
#changed bar colors for all weather types   
  geom_bar(fill = "navyblue") +
#changed the label for x-axis: Weather 
  labs(x = "Weather",
#changed the label for y-axis: Frequency of Refilling Water Bottle        
       y = "Frequency of Refilling Water Bottle",
#added a title to the bar chart        
       title = "Frequency of Refilling Water Bottle Based on Weather",
#added a subtitle
subtitle = "27 May 2026") +
#used a different theme from default  
  theme_minimal() +
#removed panel background  
  theme(panel.background = element_blank()) + 
#removed major and minor gridlines   
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) 

```

```{r}

#base layer: ggplot
ggplot(data = personal_clean, 
#x-axis: temperature (Fahrenheit)
       mapping = aes(x = temperature_f)) +
#develops a scatterplot for frequency response variables, which can have colors
  geom_point(stat = "count", color = "darkred", size = 1) + 
#relabel x- and y-axes, added a subtitle
  labs(x = "Temperature (°F)",
       y = "Frequency of Refilling Water Bottle",
       subtitle = "27 May 2026") + 
#changed theme from default  
  theme_minimal()+
#removed panel background  
  theme(panel.background = element_blank()) + 
#removed major and minor gridlines   
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) 

```

## b.



# Problem 3. Affective visualization

## a.



## b.



## c.



## d.



## e.



# Problem 4. Statistical critique

## a.



## b.



## c.



## d.


